Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2022-12-01 analysis/TAS_memo/memo_TAS_reviewers.Rmd
Update memo_TAS_reviewers.Rmd

Second, the phrase "should be reprimanded in the future (in some cases for no reason)" seems to be premised either on worries that our approach (1) requires trial judges to infer present illegal strike bias (i.e., find _Batson_ violations ) _solely_ based on strikes in past cases; and (2) will cause judges to be more likely to find _Batson_ violations when in fact no strike bias exists. Neither is the case, and we have revised the paper to make this more clear.

@@

Rather, as Figure 1 now depicts, our approach helps attorneys evaluate the inference of illegal bias in the present case based on an attorney's strikes in the present case and past cases, i.e., _before_ accounting for other evidence relevant to illegal strike bias in the present case (the _Batson_ violation itself). For that task, our approach assigns at most equal weight to past strikes and their strikes in the present case (i.e., when $\alpha = 1$). And if there is incompatibility betwen the historical strike data and the current strike data ($b_{hist} \neq b_{curr}$) for reasons identified on page X, lines Y, that inference of bias will be sensitive to the choice of $\alpha$, and therefore that inference of bias can and should be taken as weaker for that reason.

@@

@@

As explained above, the law of _Batson_ permits the inference of present strike bias based on past strikes in previous trials. The possible incompatability of $b_hist$ and $b_curr$ is thus already an issue when using past strike data to infer bias. Our contribution is to bring that issue to foreground and to advise assessing how sensitive the inference of bias is to the choice of $\alpha$ values by choosing multiple $\alpha$ values.

@@

Just focus on one in the paper, rest in the appendix. It is a decision about how conservative you want to be. We have removed the discussion of the different CIs and foculs on the 95% CI for the manuscript. We note that this does not necessarily fixed the type I error, because this coverage probability refers to the true bias parameter (that is fixed in our simulations), and we are interested in "bias detection", i.e. whether the CI includes 0 or not. We also, present sensitivity analysis for the selection of the nominal HPD level. In practice, this needs to be chosen based on how conservative the lawyer bringing up a _Batson_ challenge behaves.

@@

[REPLY]

@@

@@

@@

>4. In Section 3, the authors state that "the anticipated degree of incompatibility should inuence which value of $\alpha$ to select." As a follow-up of comment 2, if the assumption of compatibility is correct, setting $\alpha$ = 1 does lead to better power without sacrificing the Type I error. However, based on results in Figures 1 to 3, if this assumption is wrong, choosing a large $\alpha$ leads to either low power (bias detection rate) or large Type I error, depending on the bias in historical data. In the absence of such knowledge of compatibility, how does one choose the value of $\alpha$? In addition, to get around this issue, is it reasonable to treat $\alpha$ itself random and introduce a hyper prior?

>>>>

Second, the phrase "should be reprimanded in the future (in some cases for no reason)" seems to be premised either on worries that our approach (1) requires trial judges to infer present illegal strike bias (i.e., find _Batson_ violations ) _solely_ based on strikes in past cases; and (2) will cause judges to be more likely to find _Batson_ violations when in fact no strike bias exists. Neither is the case, and we have revised the paper to make this more clear. Places where these clarifications have been made are highlighted in the paragraphs that follow.

@@

Rather, as the new Figure 1 depicts, our approach helps attorneys evaluate the inference of illegal bias in the present case based on an attorney's strikes in the present case and past cases, i.e., _before_ accounting for other evidence relevant to illegal strike bias in the present case (the _Batson_ violation itself). For that task, our approach assigns at most equal weight to past strikes and their strikes in the present case (i.e., when $\alpha = 1$). And if there is incompatibility betwen the historical strike data and the current strike data ($b_{hist} \neq b_{curr}$) for reasons identified on page X, lines Y, that inference of bias will be sensitive to the choice of $\alpha$, and therefore that inference of bias can and should be taken as weaker for that reason.

@@

We hope that these modifications clarify the placement of our model in the steps from a sequence of peremptory challenges to a _Batson_ challenge actually being brought, and upheld by the trial judge. We we have also more clearly outlined the the precedence for the use of historical strike data in the _Batson_ challenge process.

@@

As explained above, the law of _Batson_ permits the inference of present strike bias based on past strikes in previous trials. The possible incompatability of $b_hist$ and $b_curr$ is thus already an issue when using past strike data to infer bias. Our contribution is to bring that issue to foreground and to advise assessing how sensitive the inference of bias is to the choice of $\alpha$ values by choosing multiple $\alpha$ values. [? We have also included the citation from reviewer 2]

@@

For the sake of clarity, we now focus on one in the paper, and have placed the results based on alternative CIs in the appendix. It is a decision about how conservative you want to be. We have removed the discussion of the different CIs and foculs on the 95% CI for the manuscript. We note that this does not necessarily fixed the type I error, because this coverage probability refers to the true bias parameter (that is fixed in our simulations), and we are interested in "bias detection", i.e. whether the CI includes 0 or not. We also, present sensitivity analysis for the selection of the nominal HPD level. In practice, this needs to be chosen based on how conservative the lawyer bringing up a _Batson_ challenge behaves.

@@

This comment is not clear to us. The real data analyses are based on real _historical_ trial data from the stae of Connecticut (see lines x to y). Users of the app have the option to run the models excluding any historical trial data (as suggested by the comment) by not selecting the name of a prosecutor or defense attorney. In this case the normal(0,2) prior is used for the bias parameter. Similarly, the app gives the user the option of altering the weight the historical strike data (if selected) by adjusting alpha (power prior parameter). Thus the user can: 1) choose not to use historical data, 2) use historical data with equal weight as the current trial (alpha = 1), or 3) down-weight the historical trial data.

@@

[INSERT] We agree. Although, unfortunately this additional information (age, etc) is not routinely collected by the courts, and therefore is difficult to consider in the context of our model. We have added a sentence (line x) to the discussion that highlights this point by calling for the need to reassess how, and more importantly what, information is collected on prospective jurors, so that models like ours can be used to my the process more fair and equittable.

@@

[INSERT] The reviewer is correct on 2 counts: 1 the bias detection rate can be as high as 100% when true bias is zero, and 2 power to detect bias is low when there is low bias in historical trials.  1. these situations occur in the presense of compatibility whih we highlight in the manuscript as arising for various reasons (see lines x), and 2. The lack of power is low due the limited number of strikes, which is why the use of compatible historical information can increase this power. This is the purpose of power prior: adjust the weight of historical data (and therefore type I error) by adjusting alpha. In reality, users of the model will not know the values of bias in the current and historical data, and so will need to rely on ... to decide what value of alpha to choose, and as we have said above, ultimately whether or not to bring a Batson challenge where they will need to discolse their use of alpha.

@@

[INSERT] Will get Eric to check this (still no response from Xiaomeng)

>4. In Section 3, the authors state that "the anticipated degree of incompatibility should influence which value of $\alpha$ to select." As a follow-up of comment 2, if the assumption of compatibility is correct, setting $\alpha$ = 1 does lead to better power without sacrificing the Type I error. However, based on results in Figures 1 to 3, if this assumption is wrong, choosing a large $\alpha$ leads to either low power (bias detection rate) or large Type I error, depending on the bias in historical data. In the absence of such knowledge of compatibility, how does one choose the value of $\alpha$? In addition, to get around this issue, is it reasonable to treat $\alpha$ itself random and introduce a hyper prior?
